Generation of TypeScript Declaration Files from JavaScript Code

### 2020-01-06 ecoop-2020/main.tex
rewrites and questions

In the next section, we examine each step in the generation process in
detail and refer back to this example for concreteness.

@@

The examples and the code base of the library are instrumented with
Jalangi \cite{DBLP:conf/sigsoft/SenKBG13} to gather data flow
information and type information at runtime. Jalangi is a configurable
framework for dynamic analysis of JavaScript. It provides several
analysis modules that were extended as needed to retrieve the required
run-time information, which is then saved in a JSON file.

@@

\begin{lstlisting}[
    caption={\textbf{dts-generate usage} - Example of how to generate a declaration file for module \texttt{abs}.},
    language=bash,
	label=code:dts-generate-example,
    float,
    abovecaptionskip=-\medskipamount
]
$ ./dts-generate abs
$ cat output/abs/index.d.ts 
export = Abs;

declare function Abs(input: string): string;
\end{lstlisting}

@@

We considered option~\ref{item:2} under the assumption that most
libraries would come with test cases. However, there is no standard
for testing JavaScript code so that test cases were difficult to reap
from the \NPM{} packages: they used different directory structures,
different (or no) testing tools, or some packages did not have tests
at all.

In the end, option~\ref{item:3} was the most viable even though there
is no standard for documentation, either. However, almost all repositories
contain README files where the library authors briefly describe what
the code does, which problem it solves, how to install the
application, how to build the code, etc. It is very common that
developers provide code examples in the readme files to show how the
library works and how to use it. This observation holds in particular
for \NPM{} packages, which are generally created to solve a specific
problem of JavaScript development.

Obtaining code examples for a specific \NPM{} package is done in three steps.
\begin{itemize}
\item Obtain the repository url from the package: The command
  \texttt{npm view <PACKAGE> repository.url} can be used for
  retrieving the url of the package's repository.

\item Retrieve the README file from the repository.

\item Extract the code examples from the README file: README files are
  written using Markdown\footnote{\url{https://www.markdownguide.org}}, a
  popular markup language. In a README file, it is customary to write
  code examples within code blocks labeled with the programming
  language, so that syntax highlighting is done correctly. Hence, we
  retrieve the code examples from code blocks labeled \texttt{js} or
  \texttt{javascript}, which both stand for JavaScript in
  Markdown.
\end{itemize}

@@

Obtaining the code fragments from the examples provided in the
README files of the repository proved to be an
appropriate and pragmatic way of extracting the developer's
intention and providing an useful initial code base with meaningful
examples, thus avoiding possible cold start problems.

@@

The dynamic analysis framework used for gathering this kind of
information is Jalangi. The configurable analysis modules enable
programming custom callbacks that get triggered with virtually any
JavaScript event. Our instrumentation observes the following events: 
\begin{itemize}
  \item Binary operations, like \lstinline{==}, \lstinline{+} or
    \lstinline{===}.  
  \item Variable declaration.
  \item Function, method, or constructor invocation.
  \item Access to an object's property.
  \item Unary operations, like \lstinline{!} or \lstinline{typeof}.
\end{itemize}

@@

\subsection{TypeScript Declaration File Generation}
\label{sec:typescr-decl-file}
\paragraph*{Overview}
The actual generation of the declaration file is the next step in the
pipeline after gathering the run-time information, as shown in
\figref{fig:tsd_generation_method_block_diagram}. It is a lightweight,
simple and fast application, which does not interact with the actual
JavaScript module at run-time. Instead, it uses the JSON output file
containing the run-time information and generates a TypeScript
declaration file which is use ready to be used in a TypeScript
project. The tool itself is written in TypeScript and runs in a
Docker container in \NodeJS.

\paragraph*{Templates}
TypeScript provides templates for writing declaration files\footnote{https://www.typescriptlang.org/docs/handbook/declaration-files/templates.html} and each template corresponds to a different way of exporting a JavaScript module. The tool uses different fields from the run-time information to detect how the module is being used in order to choose the right templates accordingly. The implemented templates are \lstinline{module}, \lstinline{module-class} and \lstinline{module-function}.

\paragraph*{Interfaces}
Finally, interfaces are created by exploring \lstinline{getField} and \lstinline{methodCall} interactions from the run-time information. The code will gather the interactions for a specific argument and build the interface by incrementally adding new properties. Interactions within the \lstinline{followingInteractions} field are recursively traversed, building a new interface in each recursion level.

\section{Evaluation}
\label{sec:dts-generate-evaluation}
After generating a declaration file for a published \NPM{} package, it is necessary to evaluate the quality of it. It was decided to compare the generated declaration file against the one uploaded to the DefinitelyTyped repository for the same module, as shown in \figref{fig:evaluation-diagram}.

@@

\paragraph*{Parsing}
Before comparison, declaration files need to be parsed so that both of them share the same structure. It was achieved by using the TypeScript Compiler API, a library developed by Microsoft that allows to traverse the Abstract Syntax Tree in an easy and intuitive way \cite{typescript-compiler-api}.

The parsing consists in creating a structure where declared \lstinline{interfaces}, \lstinline{functions}, \lstinline{classes} and \lstinline{namespaces} are stored separately. Function arguments are correctly described, identifying complex types like union types or callbacks. Optional parameters are also identified. For \lstinline{classes}, a distinction between the constructor and methods is made. Finally, syntax and semantic errors are also checked by the TypeScript Compiler API.

The tool is called \texttt{parse-dts} and is naturally written in TypeScript. It also runs in \NodeJS{} in its corresponding Docker container.

\paragraph*{Comparator}
An independent tool will compare two parsed declaration files. As described in \figref{fig:evaluation-diagram}, the comparator will compare the generated declaration file against the corresponding file in the DefinitelyTyped repository.

It was discovered that the implementation was easier when focusing on each template independently. For this implementation, only the \lstinline{module-function} was considered for the comparison.

The following criteria were applied:
\begin{itemize}
  \item Number of declared functions: Checks the number of declared and exported functions for each of both declaration files.
  \item Name of declared functions: Checks whether both of the declaration files declared a function with the same name.
  \item Number of parameters: Checks the number of parameters of the declared functions. This is checked for optional and non-optional parameters.
  \item Interfaces: Checks the number of declared interfaces and the fields within those interfaces.
  \item Errors: Indicates whether there are errors in the declaration files.
\end{itemize}

@@

@@

Figure \ref{fig:experiments-overall-funnel} shows that a declaration file was generated for 244 modules out of 6029 modules. Samples of the generated declaration files for templates \lstinline{module}, \lstinline{module-class} and \lstinline{module-function} are presented in \secref{sec:experiments-declaration-files-generation}.

@@

\subsection{Code Examples}
Retrieving the code examples from the JavaScript libraries' repositories proved to be a pragmatic way of capturing the types. However, as shown in \figref{fig:experiments-overall-funnel}, working code examples for only 2260 modules could be retrieved. The process of getting a valid code example for a module is divided in 4 blocks:
\begin{itemize}
	\item Extracting repositories url.
	\item Extracting readme files.
	\item Extracting code examples from README files.
	\item Executing code examples and discarding failing ones.
\end{itemize}

@@

\paragraph*{Repositories URL}
The url of the repositories could be retrieved for only 4974 modules. More than 1000 modules do not have the repository entry in their corresponding \texttt{package.json} files. Therefore, the \texttt{npm view <module> repository.url} command returns an empty value. This is even happening for important modules like \texttt{ace}.

\paragraph*{Readme Files}
700 modules simply do not have a readme file in their repositories. The implementation does contemplate, however, different naming conventions like \texttt{readme.md} or \texttt{README.md}.

\paragraph*{Code Examples Extraction}
The 50\% loss is mainly explained because developers did not wrap their code around a block using the \texttt{javascript} or \texttt{js} tags. Counting with code examples for 2200 modules was considered to be enough for evaluating the generation of declaration files.

\paragraph*{Code Examples Execution}
The 2260 extracted code examples were executed by installing the required packages and running the code as a \NodeJS{} application. Working and functional code examples could only be extracted for 1448 modules. 812 modules did not run correctly and were discarded. Some failing samples were analyzed and there were mainly two reasons for the failure:
\begin{enumerate}
	\item The code example had been properly extracted but the code itself was not working. It was executing the library in an unsupported way, hence the error at run-time.
	\item The extracted code example was not intended to be executed or it was not even valid JavaScript code.
\end{enumerate}

@@

The following section exhibits some samples of the 244 generated declaration files. It shows some results for each of the implemented templates: \lstinline{module}, \lstinline{module-function} and \lstinline{module-class}.

\figref{fig:experiments-results-module-function} shows the generated declaration files for simple modules like \texttt{abs}, \texttt{dirname-regex} and \texttt{escape-html}. All of them were generated using the \texttt{module-function} template. There are no differences between the generated files and the corresponding declaration files uploaded to DefinitelyTyped.

@@

export = escapeHTML;
        \end{lstlisting}
        \caption{escape-html/index.d.ts - DefinitelyTyped}
        \label{fig:experiments-results-module-function}
      \end{subfigure}

    \caption{\textbf{Module-function results} - Results are shown for modules \texttt{abs}, \texttt{dirname-regex}, \texttt{escape-html}. On the left side the generated declaration file with \lstinline{dts-generate}. On the right side the corresponding file in the DefinitelyTyped repository. Functions are correctly detected and input types are accurately inferred. Both files are parsed for comparison, as explained in \secref{sec:dts-generate-evaluation}. Therefore, subtle differences in the syntax between both files are not important.}
\end{figure}

Templates of type \lstinline{module-class} are shown for modules \texttt{flake-idgen}, \texttt{route-parser} and \texttt{timer-machine} in \figref{fig:experiments-results-module-class-flake-idgen} and \figref{fig:experiments-results-module-class-timer-machine}, respectively. Properties of interfaces and class methods are correctly generated. Optional parameters are not detected, as it was not considered for the implementation. Finally, \lstinline{module} template is presented for \texttt{is-uuid} module in \figref{fig:experiments-results-module-is-uuid}.

@@

    \caption{\textbf{Module-class results | flake-idgen} - Parameters of interface \texttt{ConstructorOptions} are correctly detected. Name of interface differs since it is automatically generated based on the name of the argument variable. Optional properties were not implemented, hence the \lstinline{undefined} type for some properties. Analogously, callback \texttt{cb} is inferred as \lstinline{undefined}.}
    \label{fig:experiments-results-module-class-flake-idgen}
\end{figure}

@@

    \caption{\textbf{Module-class results | timer-machine} - Parameter \lstinline{started} is inferred as \lstinline{undefined} instead of marking it as optional. Methods that were not executed do not appear in the generated declaration file.}
    \label{fig:experiments-results-module-class-timer-machine}
\end{figure}

@@

    \caption{\textbf{Module results | is-uuid} - Methods that were not executed are not included in the declaration file.}
    \label{fig:experiments-results-module-is-uuid}
\end{figure}

It is worth mentioning that for some libraries the declaration file in DefinitelyTyped was not correct. For example, for \texttt{datadog-metrics}, some properties of an interface were included in the generated declaration file but they were not present in the one in DefinitelyTyped. However, as shown in \figref{fig:experiments-results-module-datadog-metrics}, the properties are indeed used in the source code and should be included.

\begin{figure}[tp]
    \centering
    \begin{subfigure}{0.48\linewidth}
      \begin{lstlisting}[language=TypeScript]
export interface I__opts {
    'aggregator': undefined;
    'defaultTags': Array<any>;
    'reporter': undefined;
    'apiKey': string;
    'appKey': undefined;
    'agent': undefined;
    'host': string;
    'prefix': string;
    'flushIntervalSeconds': number;
}

@@

    \caption{\textbf{Missing properties | datadog-metrics} - Properties \texttt{aggregator} and \texttt{reporter} are not in the DefinitelyTyped version, but they appear in the generated declaration file. However, they are indeed used by the library, as exposed in lines 2 and 3 of the library's source code shown in \protect\subref{fig:subfloat-datadog-logger}.}
    \label{fig:experiments-results-module-datadog-metrics}
\end{figure}

\subsection{Evaluation}
As shown in \figref{fig:experiments-typescript-templates-distribution-definitely-typed}, 20\% of the declaration files in DefinitelyTyped are written using the \lstinline{module-function}. However, 57\% of the 244 generated declaration files are written with the \lstinline{module-function} template. Additionally, the complexity of evaluating declaration files written with the \lstinline{module-function} is considerably lower than for other templates. The evaluation for templates \lstinline{module-class} and \lstinline{module} was not implemented.

@@

@@

	\caption[TypeScript templates distribution | Generated \& DefinitelyTyped]{\textbf{TypeScript templates distribution | Generated \& DefinitelyTyped} - Out of a total of 6029, 72\% of the modules uploaded to the DefinitelyTyped repository use the \lstinline{module} template and only 20\% use the \lstinline{module-function} one. However, 57\% of the 244 generated declaration files use the \lstinline{module-function} template.
	}

>>>>

After filling in some background information on declaration files,
Section~\ref{sec:gener-typescr-decl} examines each step in the
generation process in detail and refer back to this example for
concreteness.

\section{TypeScript Declaration Files}
\label{sec:typescr-decl-files}

The declaration file shown in
\figref{fig:motivating-example-route-parser-vscode} describes a
package with a single exported class \lstinline{Route}. The
description of the class comprises the type signature of the
constructor (line~4) and the methods \lstinline{match} and
\lstinline{reverse} (lines~5 and~6).

!!!TODO: explain meaning of namespace.

This file is an instance of one of the standard templates for writing
declaration
files\footnote{https://www.typescriptlang.org/docs/handbook/declaration-files/templates.html}:
\textbf{module}, \textbf{module-class} and
\textbf{module-function}. 
Each template corresponds to a different way of describing the exports
of a JavaScript library. Choosing the template depends on the
particular organization of the underlying JavaScript library:
\begin{description}
\item[module] several exported functions,
\item[module-class] a class-like structure,
\item[module-function] exactly one exported functions.
\end{description}

% $ ./dts-generate abs
% $ cat output/abs/index.d.ts 
\begin{lstlisting}[
    caption={Example for template \textbf{module-function}},
    language=TypeScript,
	label=code:dts-generate-example,
    float,
    abovecaptionskip=-\medskipamount
]
export = Abs;

declare function Abs(input: string): string;
\end{lstlisting}
!!!TODO: example for simple \textbf{module}

The \texttt{route-parser} library is an instance of a
\textbf{module-class}. The \texttt{abs} library provides a single
function \lstinline{Abs} so it uses template \textbf{module-function}
as shown in \ref{code:dts-generate-example}.

@@

The examples and the code base of the library are instrumented with
Jalangi \cite{DBLP:conf/sigsoft/SenKBG13} to gather data flow
information and type information at runtime. Jalangi is a configurable
framework for dynamic analysis of JavaScript. It provides several
analysis modules that we extended as needed to retrieve the required
run-time information, which is then saved in a JSON file.

@@

@@

We considered option~\ref{item:2} under the assumption that most
libraries would come with test cases. However, there is no standard
for testing JavaScript code so that test cases were difficult to reap
from the \NPM{} packages: they use different directory structures,
employ differing (or no) testing tools, or do not have tests
at all.

In the end, option~\ref{item:3} was the most viable even though there
is no standard for documentation, either. However, almost all repositories
contain README files where the library authors briefly describe in prose what
the code does, which problem it solves, how to install the
application, how to build the code, etc. It is very common that
developers provide code examples in the README files to show how the
library works and how to use it. This observation holds in particular
for \NPM{} packages, which are generally created to solve a specific
problem of JavaScript development.

Obtaining code examples for a specific \NPM{} package is done in three steps.
\begin{itemize}
\item Obtain the  repository's URL with the command
  \texttt{npm view <PACKAGE> repository.url}

\item Retrieve the README file from the top-level directory of the repository.

\item Extract the code examples from the README file. To this end,
  observe that README files are
  written using Markdown\footnote{\url{https://www.markdownguide.org}}, a
  popular markup language. In such a file, it is customary to write
  code examples in code blocks labeled with the programming
  language, so that syntax highlighting is done correctly. Hence, we
  retrieve the code examples from code blocks labeled \texttt{js} or
  \texttt{javascript}, which both stand for JavaScript in
  Markdown.
\end{itemize}

@@

Obtaining the code fragments from the examples provided in the
README files of the repository proved to be an
appropriate and pragmatic way of extracting the developer's
intention and providing a useful initial code base with meaningful
examples, thus avoiding possible cold start problems.

@@

The dynamic analysis framework Jalangi is used for gathering this kind of
information. The configurable analysis modules enable
programming custom callbacks that get triggered with virtually any
JavaScript event. Our instrumentation observes the following events: 
\begin{itemize}
  \item Binary operations, like \lstinline{==}, \lstinline{+} or
    \lstinline{===}.  
  \item Variable declaration.
  \item Function, method, or constructor invocation.
  \item Access to an object's property.
  \item Unary operations, like \lstinline{!} or \lstinline{typeof}.
\end{itemize}

@@

\subsection{TypeScript Declaration File Generation}
\label{sec:typescr-decl-file}
%\paragraph*{Overview}

The next step in the pipeline after gathering the run-time information is
the actual generation of the declaration file (cf.\
\figref{fig:tsd_generation_method_block_diagram}). It is a lightweight,
simple and fast application, which solely relies on the run-time
information gathered in the previous step.
% The tool itself is written in TypeScript and runs in a
% Docker container in \NodeJS.

!!!TODO too vague: more information --- how do you choose between the templates?

We use different fields (which?) from the run-time information to detect
how the module is used to choose the most suitable template. The implemented
templates are \textbf{module}, \textbf{module-class} and
\textbf{module-function}.

% \paragraph*{Interfaces}
Finally, interfaces are created by exploring \texttt{getField} and
\texttt{methodCall} interactions from the run-time information. We
gather the interactions for a specific argument and build the
interface by incrementally adding new properties. Interactions within
the \texttt{followingInteractions} field are recursively traversed,
building a new interface in each recursion level.

!!!TODO: insufficient explanation of \texttt{followingInteractions}
--- needs to be mentioned in the previous step

\section{Evaluation}
\label{sec:dts-generate-evaluation}
After generating a declaration file for an \NPM{} package, we need to
evaluate its quality. To this end, we compare the generated
declaration file against the one uploaded to the DefinitelyTyped
repository for the same module, as shown in
\figref{fig:evaluation-diagram}. While this approach does not provide
an absolute measure of quality, it gives us at least an indication of
the pragmatics of \texttt{dts-generate}: The files on DefinitelyTyped
are perceived to be useful by the community. If the accuracy of the
generated files is comparable with the accuracy of the files on
DefinitelyTyped, then \texttt{dts-generate} is a viable alternative to
manually creating a definition file.

!!!TODO: check how many files on DT are just generated by
\texttt{dts-gen} (e.g., the TypeScript tool)

@@

To this end, we need a means of comparing two declaration files. Of
course, we are not interested in a textual comparison, but in a
comparison of the structures described by the files. Technically, our
comparison of two structures A and B yields one of five outcomes:
\begin{itemize}
\item A and B are equivalent;
\item A is more general than B;
\item B is more general than A;
\item A and B have a common generalization (i.e., they are
  compatible);
\item none of the above holds (i.e., A and B are incompatible).
\end{itemize}

The first step is to parse declaration files, which can be
achieved by using the TypeScript Compiler API, a library traverse the
Abstract Syntax Tree of a TypeScript program
\cite{typescript-compiler-api}. 
The step also performs a sanity check of the generated declaration
files as it rejects files with syntactic or semantics errors.
The output of the parser is a structure where declared
{interfaces}, {functions}, {classes} and
{namespaces} are stored separately. Function arguments are
correctly described, identifying complex types like union types or
callbacks. Optional parameters are also identified. For
{classes}, a distinction between the constructor and methods
is made.

% The tool is called \texttt{parse-dts} and is naturally written in
% TypeScript. It also runs in \NodeJS{} in its corresponding Docker
% container.

It remains to compare the parsed representations of two declaration
files.
% It was discovered that the implementation was easier when
% focusing on each template independently.
As an example, we discuss the implementation of comparison for
the \textbf{module-function} template.

The following criteria were applied:
\begin{itemize}
\item Number of declared functions: Checks the number of declared
  and exported functions for each of both declaration files. 
\item Name of declared functions: Checks whether both of the
  declaration files declared a function with the same name. 
\item Number of parameters: Checks the number of parameters of the
  declared functions. This is checked for optional and non-optional
  parameters.

  !!!TODO: what about parameter names? what about object
  types/interface types of parameters?
\item Interfaces: Checks the number of declared interfaces and the
  fields within those interfaces. 
\item Errors: Indicates whether there are errors in the declaration files.
\end{itemize}

!!!TODO: too informal. we need to define an ordering on interfaces etc.

@@

!!!TODO: first we need to say, what we want to achieve with the
experimental evaluation

@@

Figure \ref{fig:experiments-overall-funnel} shows that a declaration
file was generated for 244 modules out of 6029 modules. Samples of the
generated declaration files for templates \textbf{module},
\textbf{module-class} and \textbf{module-function} are presented in
\secref{sec:experiments-declaration-files-generation}.

@@

\subsection{Code Examples}
Retrieving the code examples for the JavaScript libraries proved to be
a pragmatic way of driving the type gathering at run time. However, as
shown in \figref{fig:experiments-overall-funnel}, it was only possible
to obtain working code examples for 2260 packages. The
process of getting a valid code example for a module is divided in four
stages: 
\begin{itemize}
\item Extracting repository URL.
\item Extracting README file.
\item Extracting code examples from README files.
\item Executing code examples and discarding failing ones.
\end{itemize}

@@

\paragraph*{Repository URL}
The URL of the source repository could be retrieved for only 4974
packages. More than 1000 packages on \NPM{} do not have a repository
entry in their corresponding \texttt{package.json} file. Therefore, the
\texttt{npm view <module> repository.url} command returns no
value. Even important modules like \texttt{ace} provide no repository URL.

\paragraph*{README Files}
700 packages do not have a README file in their repository, although
the implementation checks for several naming conventions like
\texttt{readme.md} or \texttt{README.md}.

\paragraph*{Extraction of Code Examples}
In this step, we loose another 50\% of modules! This loss is mainly
explained because some developers do not wrap their code in a block
using the \texttt{javascript} or \texttt{js} tags. However, as we are
still left with code examples for 2200 modules, we did not further
look into code extraction as this number was considered sufficient for
evaluating the generation of declaration files.

\paragraph*{Execution of Code Examples}
We executed the remaining 2260 extracted code by installing the
required packages and running the code as a \NodeJS{}
application. Unfortunately, the code examples only worked for 1448
modules. 812 modules did not run correctly and had to be
discarded. Some failing samples were analyzed and there were mainly
two reasons for the failure: 
\begin{itemize}
\item The code fragment had been properly extracted but the code was
  not faulty. It was executing the library in an unsupported
  (obsolete?) way, which lead to a run-time error.
\item The extracted code fragment was not intended to be executed
  and/or it was not even valid JavaScript code. 
\end{itemize}

!!!TODO: the remaining two stages ``run-time information'' and
``generated declaration files'' need to be explained

@@

This section exhibits some samples of the 244 generated
declaration files. It shows some results for each of the implemented
templates: \textbf{module}, \textbf{module-function} and
\textbf{module-class}.

\figref{fig:experiments-results-module-function} shows the generated
declaration files for simple modules like \texttt{abs},
\texttt{dirname-regex}, and \texttt{escape-html}. All of them were
generated using the \textbf{module-function} template. There are no
differences between the generated files and the corresponding
declaration files on DefinitelyTyped.

The left side of the figure shows the generated declaration file with
\lstinline{dts-generate}, the right side shows the corresponding file
in the DefinitelyTyped repository. Functions are correctly detected; input
and output types are accurately inferred.

!!!TODO: explain, what about
\begin{itemize}
\item uppercase/lowercase in dirname-regex?
\item extra namespace in escape-html?
\end{itemize}

@@

export = escapeHTML;
        \end{lstlisting}
        \caption{escape-html/index.d.ts - DefinitelyTyped}
      \end{subfigure}

    \caption{Results for \textbf{module-function} template}
    \label{fig:experiments-results-module-function}
  \end{figure}

Templates of type \textbf{module-class} are shown for modules
\texttt{flake-idgen}, \texttt{route-parser}, and \texttt{timer-machine}
in \figref{fig:experiments-results-module-class-flake-idgen} and
\figref{fig:experiments-results-module-class-timer-machine},
respectively. Properties of interfaces and class methods are correctly
generated. Optional parameters are not detected, as it was not
considered for the implementation.

For \texttt{flake-idgen},  the parameters of interface
\texttt{ConstructorOptions} are correctly detected. The name of the
interface differs because it is automatically generated based on the
name of the argument variable. Optional properties were not
implemented, which explains the type \lstinline{undefined} for some
properties. Analogously, the callback \texttt{cb} is inferred as
\lstinline{undefined}.

!!!TODO:
\begin{itemize}
\item why is the callback undefined? is it due to the example? (what does it
  look like?) why is it called \texttt{cb}?
\item the export / class is named differently; doesn't that matter?
\end{itemize}

For \texttt{timer-machine},  the parameter \lstinline{started} is
inferred as \lstinline{undefined} instead of marking it as
optional. Methods that were not executed cannot appear in the
generated declaration file. For that reason, a type corresponding to
\lstinline{TimerEvent} is not generated, either.

@@

    \caption{Results for \textbf{module-class} module \texttt{flake-idgen}}
    \label{fig:experiments-results-module-class-flake-idgen}
\end{figure}

@@

    \caption{Results for \textbf{module-class} module \texttt{timer-machine}}
    \label{fig:experiments-results-module-class-timer-machine}
\end{figure}

@@

    \caption{Results for \textbf{module}  module \texttt{is-uuid}}
    \label{fig:experiments-results-module-is-uuid}
\end{figure}

Finally, a sample for the \textbf{module} template
for the \texttt{is-uuid} module is shown in
\figref{fig:experiments-results-module-is-uuid}.
Again, methods that were not executed by the extracted examples are
not included in the declaration file.

It is worth mentioning that for some libraries the declaration file in
DefinitelyTyped was not correct. For example, for the module
\texttt{datadog-metrics}, some properties of an interface were
included in the generated declaration file but they were not present
in the declaration presented in DefinitelyTyped. However, as shown in
\figref{fig:experiments-results-module-datadog-metrics}, the
properties \texttt{aggregator} and \texttt{reporter} are not present
in the DefinitelyTyped version, but they appear in the generated
declaration file. However, they are indeed used by the library, as
exposed in lines~2 and~3 of the library's source code shown in
\figref{fig:subfloat-datadog-logger}. 
\begin{figure}[tp]
    \centering
    \begin{subfigure}{0.48\linewidth}
      \begin{lstlisting}[language=TypeScript]
export interface I__opts {
    'aggregator': undefined;
    'defaultTags': Array<any>;
    'reporter': undefined;
    'apiKey': string;
    'appKey': undefined;
    'agent': undefined;
    'host': string;
    'prefix': string;
    'flushIntervalSeconds': number;
}

@@

    \caption{Missing properties for module \texttt{datadog-metrics}}
    \label{fig:experiments-results-module-datadog-metrics}
\end{figure}

\subsection{Evaluation}

!!!TODO: explain what's going on in this subsection

!!!TODO: what's the baseline for the percentages in the paragraph, the
244 modules considered or all of DT? Ah, you say that in the figure;
but then the figure is comparing apples with pears (percentages for
different baselines!).

As shown in
\figref{fig:experiments-typescript-templates-distribution-definitely-typed},
20\% of the declaration files in DefinitelyTyped are written using the
\textbf{module-function}. However, 57\% of the 244 generated
declaration files are written with the \textbf{module-function}
template. Additionally, the complexity of evaluating declaration files
written with the \textbf{module-function} is considerably lower than
for other templates. The evaluation for templates
\textbf{module-class} and \textbf{module} was not implemented.

@@

!!!TODO: that's a bit lame as a result. See above for a more
meaningful output.

@@

	\caption[TypeScript templates distribution | Generated \& DefinitelyTyped]{\textbf{TypeScript templates distribution | Generated \& DefinitelyTyped} - Out of a total of 6029, 72\% of the modules uploaded to the DefinitelyTyped repository use the \textbf{module} template and only 20\% use the \textbf{module-function} one. However, 57\% of the 244 generated declaration files use the \textbf{module-function} template.
	}
